---
output: 
  html_document: 
    keep_md: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```
#Data Science Using R
##2 - R Programming

###0) R version used
```{r}
print(R.version.string)
```

###1) basic datatypes: input/eval/Robject/vectorlists/matrix/factors
```{r}
print(1:20)
print(1+2i)
print(c(TRUE,"a"))
x<-list(1,"a",TRUE,1+4i)
print(x)
m<-matrix(nrow=2,ncol=3)
print(m)
dim(m)
print(matrix(1:8,nrow=2,ncol=4))
num<-1:10
dim(num)<-c(5,2)
print(num)
cbind(1:3,10:12)
rbind(1:3,10:12)
x<-factor(c("yes","yes","no","yes","no"))
x
table(x)
unclass(x)
y<-factor(c("yes","yes","no","yes","no"), levels=c("yes","no"))
y
x<-c(1,2,NA,3)
is.na(x)
y<-c(1,2,NaN,NA,6)
is.na(y)
is.nan(y)
x<-data.frame(foo=1:4,bar=c(T,T,F,F))
x
print(paste("nrow:", nrow(x),", ncol:",ncol(x)))
x<-1:3
names(x)
names(x)<-c("foo","bar","norf")
names(x)
x<-list(a=1,b=2,c=3)
x
m<-matrix(1:4,nrow=2,ncol=3)
dimnames(m)<-list(c("a","b"),c("c","d","e"))
m
```

###2) basic reading: read.table/dput/dget/dump/source/
```{r}
data<-read.table("foo.txt")
data
foo<-read.table(file="test.txt",header=TRUE,sep="\n",nrows=3)
foo
# passing adata around by deparsing R object using dput
# and reading back using dget
y<-data.frame(a=1,b="a")
#dput(y)
dput(y,file="y.R") # actually creates physical y.R
new.y<-dget("y.R")
new.y
x<-"foo"
y<-data.frame(a=1, b="a")
dump(c("x","y"),file="data.R") # creates physical data.R
rm(x,y)
source("data.R")
y
x
```

###3) basic connections: file/readLines/writeLines/url
```{r}
con<-file("HMP_16s_single_cell_MDA_from_fecal_specimens-short.txt","r")
x<-readLines(con,100)
x
out<-file("test-output.txt","w")
writeLines(text = x,con = out)
close(con)
close(out)
con<-url("http://www.google.com","r")
x<-readLines(con)
head(x,1)
close(con)
```

###4) basic subsetting: vector/list/matrix/partial-matching/removing-NA-values
```{r}
library(datasets)
x<-c("a","b","c","d","e")
x[1:4]
x[x>"b"]
y<-list(foo=1:4,bar=c(0.6,0.4,0.2),baz=TRUE,bad=list(4,3,2,1))
y[[2]]
y$bar
y[["bar"]]
y[["bad"]][[3]]
m<-matrix(1:6,2,3)
m
m[1,2] #m[row,col]
m[1,] #m entire row=1
m[1,,drop=FALSE] # if drop=FALSE, the 1x1 matrix is maintained
x<-list(aardvark=1:5)
x[["aard",exact=FALSE]] # if exact=FALSE, partial matching is ok
x<-c(1,2,NA,4,NA,5)
x[!is.na(x)]
y<-c("a","b",NA,"d",NA,NA)
x[complete.cases(x,y)]
y[complete.cases(x,y)]
airquality[1:10,]
airquality[complete.cases(airquality),][1:10,]
```

###5) basic vectorized operations
```{r}
x<-1:4;y<-6:9
x*y
x<-matrix(1:4,2,2);y<-matrix(rep(10,4),2,2)
x*y # element-wise multiplication
x%*%y # true matrix multiplication
y%*%x
```

###6) basic control structures: if-else/for/while/repeat
```{r}
x<-5;y<-7
if(x>y) {
    print("x is bigger than y")
} else if(x<y) {
    print("y is bigger than x")
} else {
    print("x and y are same")
}
z<- if(x>3) {
    10
} else {
    30
}
z
for(i in 4:1) {
    print(i)
}
x<-c("d","c","b","a")
for(i in 1:4){ print(x[i]) } # 1 of 3 ways to achieve similar result
for(i in seq_along(x)){ print(x[i]) } # 1 of 3 ways to achieve similar result
for(letter in x) { print(letter) } # 1 of 3 ways to achieve similar result
x<-matrix(1:6,2,3)
x
# nested for-loop
for(i in seq_len(nrow(x))) {
    for(j in seq_len(ncol(x))) {
        print(x[i,j])
    }
}
count<-0
while(count<10) {
    print(count)
    count<-count+1
}
z<-5
while(z>=3 && z<=10) {
    print(z)
    coin<-rbinom(1,1,0.5)
    if(coin==1) { # random coin toss-up
        z<-z+1
    } else {
        z<-z-1
    }
}
x0<-1
i<-0
repeat{ #infinite while(TRUE) loop
    i<-i+1
    x1<-rnorm(1)
    if(abs(x1-x0)>6) {
        print(abs(x1-x0))
        print(i)
        break
    } else {
        x0<-x1
    }
}
for(i in 1:100) {
    if(i<=90) { # skip from 1 to 90
        next
    }
    print(i)
}
```

###7) basic functions: function/...arg/args
```{r}
library(datasets)
above<-function(x,n) {
    use<-x>n
    x[use]
}
above(1:20,12)
colmeans<-function(y) {
    nc<-ncol(y)
    means<-numeric(nc)
    for(i in 1:nc){
        #means[i]<-mean(y[,i])
        means[i]<-mean(y[,i][!is.na(y[,i])])
    }
    means
}
colmeans(airquality)
myplot<-function(x,y,type="l",...){
    plot(x,y,type=type,...) # other arguments after x,y,type as ...
}
myplot(1,1,main="test")
args(paste)
paste("a","b",sep=":") # after ..., has to be explicitly named
paste("a","b",se=":") # partial match invalid
```

###8) basic scoping rules: symbol-binding/scoping-rules
```{r}
lm<-function(x) { x * x }
lm # ^ above function is called, not the lm function in stats package
# R uses lexical scoping / static scoping
# NOT dynamic scoping
f<-function(x,y) { x^2 + y/z }
z<-2;f(10,100)
f<-function(x,y) { z<-4; x^2 + y/z }
z<-2;f(10,100)
make.power<-function(n){
    pow<-function(x){
        x^n
    }
    pow
}
# function returns another function
cube<-make.power(3)
cube(3)
square<-make.power(2)
cube(2)
y<-10
f<-function(x){y<-2;y^2+g(x)}
g<-function(x){x*y}
f(3) # 4 + g(3) ==> 4 + (3*10) = 34; different scope of y
```

###9) basic dates and times
```{r}
unclass(as.Date("2016-02-09")) # days since 1970-01-01
# time is either POSIXct (large integer) or POSIXlt (list + other info)
x<-Sys.time()
p<-as.POSIXlt(x)
paste(p$hour,p$min,p$sec)
unclass(Sys.time()) # time since 1970-01-01
datestring<-c("January 10, 2012 10:40", "December 9, 2011 9:10")
strptime(datestring, "%B %d, %Y %H:%M")
as.Date("2012-03-01")-as.Date("2012-02-08")
as.POSIXct("2015-02-09 13:47:00", tz="PST8PDT")-as.POSIXct("2015-02-09 13:47:00", tz="EST5EDT")
```

###10) basic loop functions: lapply/sapply/apply/mapply/tapply/split
```{r}
# lapply: loop over list & eval a function on each element
# input: list/vector
# output: list
x<-list(a=1:5,b=rnorm(10))
lapply(x,mean)
lapply(1:4,runif,min=0,max=10) # min max arg is for runif func (random uniform variables)

# sapply: simplified result of lapply
sapply(list(a=1.5,b=rnorm(10),c=rnorm(20,1),d=rnorm(100,5)),mean)

# apply: eval a function over margins of an array
# not much faster than a loop (but simiplified in one line)
# input: array or matrix
# output: vector or matrix
m<-matrix(rnorm(200),2,5)
m
# apply sum for each row (1)
apply(m,1,sum) # ~ rowSums(m); 1 indicates row, 2 indicates column

# average matrix values in an array
a<-array(rnorm(2*2*10),c(2,2,10))
apply(a,c(1,2),mean) # ~ rowMeans(a,dims=2)

# mapply: multivariate apply of sorts -- function eval in parallel
# input: functions
# output: list/vector
mapply(rep,1:4,4:1) # ~ list(rep(1,4),rep(2,3),rep(3,2),rep(4,1))
# vectorizing a function
noise<-function(n,mean,sd,sed){set.seed(sed);rnorm(n,mean,sd)}
# list of vectors of noise(1,1,2,37), noise(2,2,2,37), et cetera
mapply(noise,1:5,1:5,2,37)
# tapply: apply function over subsets of vector
# input:
# output:
x<-c(rnorm(2),runif(2),rnorm(2,1))
x;f<-gl(3,2)
tapply(x,c(1,1,2,2,3,3),sum) # apply Factor and add sum according to such Factor
tapply(x,f,sum,simply=FALSE) # disables simplificationn
tapply(x,f,range)
# splits per Factor (f) as subvector of x
split(x,f)
lapply(split(x,f),sum) # can take the split subvector and apply sum
# splitting a data frame
library(datasets)
s<-split(airquality,airquality$Month) # split based on the Month (as Factor)
lapply(s,function(x){
    colMeans(x[,c("Ozone","Solar.R","Wind")],na.rm=TRUE)})
sapply(s,function(x){
    colMeans(x[,c("Ozone","Solar.R","Wind")],na.rm=TRUE)})
# splitting on more than one level
interaction(1:5,5:1)
split(rnorm(4),list(1:2,2:1))
split(rnorm(4),list(1:2,2:1),drop=TRUE)
```

###11) str function
```{r}
x<-rnorm(100,2,4)
summary(x)
str(x)
f<-gl(10,10)
f
summary(f)
str(f)
library(datasets)
str(airquality)
m<-matrix(rnorm(100),10,10)
str(m)
str(split(airquality,airquality$Month))
```

###12) basic simulation: random-num-gen/sim-linear-model/random-sampling
```{r}
# rnorm: random normal variates with mean/std
rnorm(10,10,1)
# dnorm: probability density with mean/std at point/vector of points
dnorm(10,10,1)
# pnorm: eval cumulative distribution function for normal distribution
pnorm(10,10,1)
# rpois: gen random Poisson variates with given rate
rpois(10,10)

x<-rnorm(10)
summary(x)

set.seed(1)
rpois(10,1)
# cumulative distribution
ppois(2,2) ## Pr(x<=2)
ppois(4,2) ## Pr(x<=4)
ppois(6,2) ## Pr(x<=6)

# linear model:
# y = ß0 + ß1x + ε 
# where ε ~ N(0.2^2)
# assume x ~ N(0.1^2), ß0 = 0.5, ß1=2
set.seed(20)
x<-rnorm(100)
e<-rnorm(100,0,2)
y<-0.5+2*x+e
summary(y)
plot(x,y)

# if x is binary
set.seed(10)
x<-rbinom(100,1,0.5)
e<-rnorm(100,0,2)
y<-0.5+2*x+e
summary(y)
plot(x,y)

# simulation of Poisson model
# Y ~ Poisson(µ)
# log µ = ß0 + ß1x
# and ß0 = 0.5 and ß1 = 0.3
set.seed(1)
x<-rnorm(100)
log.mu<-0.5+0.3*x
y<-rpois(100,exp(log.mu))
summary(y)
plot(x,y)

# random sampling
set.seed(1)
sample(1:100,4) # randomly sampling 4 in vector of 1:100
sample(letters,5)
sample(1:10) # permutation
sample(1:10,replace=TRUE) # with replacement
```